---
title: "FWS Eagle Mortality Model"
author: "Center for Conservation Innovation"
output: 
  flexdashboard::flex_dashboard:
    css: www/dashboard.css
    social: menu
    df_print: paged
    navbar:
    - {title: "CCI", align: left, href: "https://cci-dev.org"}
    - {title: "Defenders", align: left, href: "http://www.defenders.org"}
    - {title: "", icon: "fa-question-circle fa-lg", align: right, href: "mailto:esa@defenders.org?subject=Five-year reviews app"}
    - {title: "", icon: "fa-github fa-lg", align: right, href: "https://github.com/Defenders-ESC/"}
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(dplyr)
library(DT)
library(flexdashboard)
library(googlesheets)
library(ggplot2)
library(plotly)
library(shiny)
library(viridis)

load("data/app_data.RData")
source("R/helper_fxns.R")
#source("exposure.R")
#source("fatality.R")

prior <- curve(dgamma(x, shape = expose$shape,  
                      rate = 2.76),#expose$rate),
               from = 0.01,
               to = 6)

priorb <- curve(dgamma(x, shape = expose$shape,
                       rate = expose$rate),
                from = 0,
                to = 10)

collision <- curve(dbeta(x, shape1 = collide$shape, shape2 = collide$rate), 
                   from = 0, 
                   to = 0.04)

collisionb <- curve(dbeta(x, shape1 = collide$shape, shape2 = collide$rate),
                    from = 0,
                    to = 0.04)

act <- mean(Bay16$FLIGHT_MIN)/mean(Bay16$EFFORT)

tab <- select(Bay16, 
                    SITE, TURBINES, 
                    EFFORT, FLIGHT_MIN)

dt <- DT::datatable(tab,
                    fillContainer = TRUE,
                    selection = list(mode = 'single', 
                                     target = 'row'),
                    colnames = c("Site", "Turbines", "Survey Effort (hr*km3)",
                                 "Eagle Obs (min)"),
                  options = list(rownames = FALSE,
                                 pageLength = nrow(tab),
                                 dom = 'tip'
                                 )
                )%>%
  formatRound("EFFORT", 2)

# Create model dialog
introModal <- modalDialog(
  title = h2("Bayesian Eagle Mortality Prediction"),
  p("The ", tags$a(href = "https://www.fws.gov/", "U.S. Fish and Wildlife Service")," uses a Bayesian model to estimate the number of bald and golden eagles likely to be killed by proposed wind projects.
This approach combines surveys of eagle activity at proposed sites with prior information about eagle collision and activity rates across existing wind farms to estimate the likely number of fatalities."),
br(),
p("This dashboard allows you to explore how priors interact with survey data to produce predicted eagle fatalities at a set of wind energy sites. Eagle fatalities are predicted using four pieces of information obtained prior to construction:"),
tags$ol(
  tags$li("Project Specs: the number and size of turbines."),
  tags$li("Survey Area: the size of the area surveyed for eagles."),
  tags$li("Survey Hours: time spent conducting eagle surveys."),
  tags$li("Flight Time: the total time eagles were observed flying within survey areas")
),
  easyClose = TRUE
  )

# create a button to show the modal 
expand <- actionButton("show", "Click for Details")
observeEvent(input$show,
             {showModal(introModal)})

# define inputs
selectTurbines <- numericInput('n',
                               'Number of Turbines',
                               value = 100,
                               min = 0, 
                               max = 1000,
                               width = '100%')

selectHours <- numericInput('t',
                            'Hours of Operation',
                            value = 10,
                            min = 8, 
                            max = 14,
                            width = '100%')
selectHeight <- numericInput('h',
                             'Turbine Height (m)',
             value = 200,
             min = 50,
             max = 300,
             width = '100%')

selectRadius <-  numericInput('r',
                              'Rotor Radius (m)',
             value = 30,
             min = 20,
             max = 100,
             width = '100%')

plotSlider <- sliderInput("area",
            "Survey plots (#)",
            min = 0,
            max = 20,
            value = 1,
            step = 1)

timeSlider <- sliderInput("hrs",
            "Survey Hours per plot",
            min = 0,
            max = 48,
            value = 1,
            step = 0.5)

flightMin <- numericInput("min",
            "Flight Time (min)",
            value = 7,
            min = 0)

# reactive functions using inputs to define model parameters
cur_min <- reactive({input$min})

cur_effort <- reactive({input$hrs * (input$area) * 0.402})

cur_scale <- reactive({(input$t*365) * input$n * input$h/1000 * ((input$r/1000)^2) * pi})

a <- reactive({expose$shape + cur_min()})
b <- reactive({expose$rate + cur_effort()})

exposurePlot <- function(min, hours, area){
  b <- hours*area
  post <- curve(dgamma(x, min + expose$shape, b+expose$rate),
                from = 0.01, to = 6)
  obs <- min/b
  plot_ly()%>%
      add_trace(name = 'Prior',
                x = prior$x, y = prior$y,
                type = "scatter", mode = "lines", fill = "tozeroy",
                line = list(color = vir_col(3)[1]),
                fillcolor = vir_col(3)[1],
                text = ~paste("Prior Activity<br>estimate = ",
                              round(prior$x, 2),
                              sep = ""),
                hoverinfo = "text")%>%
      add_trace(name = "Survey + prior", 
                x = ~post$x, y = ~post$y,
                type = "scatter", mode = "lines", fill = "tozeroy",
                line = list(color = vir_col(3)[3]),
                fillcolor = vir_col(3)[3],
                text = ~paste("Combined Activity<br>estimate = ",
                              round(post$x, 2),
                              sep = ""),
                hoverinfo = "text")%>%
      add_trace(name = 'Observed',
                x = ~c(obs, obs), y = ~c(0,1.5),#max(c(prior$y,post$y))),
                type = "scatter", mode = "lines",
                line = list(color = 'black'),
                text = ~paste("Observed Activity<br>at site = ",
                              round(act,2),
                              sep = ""),
                hoverinfo = "text")%>%      
      layout(#title = "Eagle Exposure",
             xaxis = list(title = "Eagle Activity (min/km<sup>3</sup>*hr)",
                          range = c(0,6)),
             yaxis = list(title = "Probability Density"),
             legend = list(x = 0.7,
                           y = 1))
}
    
output$exposure <- renderPlotly({exposurePlot(input$min, input$hrs, input$area)})
```

Inputs {.sidebar data-width=400}
====================

### About

This dashboard demonstrates the behavior of the Collision Risk Model used to predict eagle fatalities at proposed wind energy facilities. Click below for more details.

```{r expand, echo = FALSE}
expand
```

### <i class="fa fa-sliders"></i> Parameters

Project specs:

```{r selector, echo = FALSE}
column(6, selectTurbines, selectHours)

column(6, selectHeight, selectRadius)
```

Survey area (FWS minimum = 1 plot ~ 201 ha)

```{r echo = FALSE}
plotSlider
```

Survey time (FWS minimum = 24 hrs)

```{r echo = FALSE}
timeSlider
```

Observed eagle flight time

```{r echo = FALSE}
flightMin 
```

`r hr()`

```{r calculate, echo = FALSE}
actionButton("calculate", "Calculate Fatalities")
helpText("Note: Calcuating predicted fatality distributions takes time.  Please allow several seconds for the graph to load.")
```

`r hr()`
===

Column 
--------------------------------------------------------------------------

### Eagle Activity

```{r exposures, echo = FALSE}
# observeEvent(input$update, {
#   obs <- isolate({cur_min()/cur_effort()})
#   post <- isolate({curve(dgamma(x, shape = a(), rate = b()), from = 0, to = 6)})
#   
#   output$exposure <- renderPlotly({
#     plot_ly()%>%
#       add_trace(name = 'Prior',
#                 x = prior$x, y = prior$y,
#                 type = "scatter", mode = "lines", fill = "tozeroy",
#                 line = list(color = vir_col(3)[1]),
#                 fillcolor = vir_col(3)[1],
#                 text = ~paste("Prior Activity<br>estimate = ",
#                               round(prior$x, 2),
#                               sep = ""),
#                 hoverinfo = "text")%>%
#       add_trace(name = "Survey + prior", 
#                 x = ~post$x, y = ~post$y,
#                 type = "scatter", mode = "lines", fill = "tozeroy",
#                 line = list(color = vir_col(3)[3]),
#                 fillcolor = vir_col(3)[3],
#                 text = ~paste("Combined Activity<br>estimate = ",
#                               round(post$x, 2),
#                               sep = ""),
#                 hoverinfo = "text")%>%
#       add_trace(name = 'Observed',
#                 x = ~c(obs, obs), y = ~c(0,1.5),#max(c(prior$y,post$y))),
#                 type = "scatter", mode = "lines",
#                 line = list(color = 'black'),
#                 text = ~paste("Observed Activity<br>at site = ",
#                               round(act,2),
#                               sep = ""),
#                 hoverinfo = "text")%>%      
#       layout(#title = "Eagle Exposure",
#              xaxis = list(title = "Eagle Activity (min/km<sup>3</sup>*hr)",
#                           range = c(0,6)),
#              yaxis = list(title = "Probability Density"),
#              legend = list(x = 0.7,
#                            y = 1))
#    })
# 
# })

plotlyOutput("exposure")
```

> Activity is the number of minutes eagles are observed flying within survey areas, per hour.

### Predicted Fatalities (_Click 'Predict Fatalities'_)

```{r fatalitiesb, echo = FALSE}
observeEvent(input$calculate,{
  #if(input$sites != ""){
    outb <- isolate({simFatal(iters = 10000, BMin = cur_min(), SmpHrKm = cur_effort(),
                             aPriExp=expose$shape, bPriExp=expose$rate,
                             aPriCPr=collide$shape, bPriCPr=collide$rate)})
    fatalityb <- isolate({density(outb$fatality*cur_scale())})
    q80b <- isolate({quantile(outb$fatality*cur_scale(), 0.8)})

    out2b <- isolate({simFatal(iters = 10000, BMin = cur_min(), SmpHrKm = cur_effort(),
                             aPriExp = 0, bPriExp = 0,
                             aPriCPr = collide$shape, bPriCPr = collide$rate)})
    fatality2b <- isolate({density(out2b$fatality*cur_scale())})
    q82b <- isolate({quantile(out2b$fatality*cur_scale(), c(0.8))})
    
output$fatalityb <- renderPlotly({
        plot_ly()%>%
          add_trace(name = "Survey + prior", 
                    x = ~fatalityb$x, y = ~fatalityb$y,
                    type = "scatter", mode = "lines",
                    fill = "tozeroy",
                    fillcolor = vir_col(3)[3],
                    line = list(color = vir_col(3)[3]),
                    text = ~paste("Probability of ", 
                                  round(fatalityb$x, 2),
                                  " fatalities<br>using prior = ",
                                  round(fatalityb$y, 2),
                                  sep = ""),
                    hoverinfo = "text")%>%
          add_trace(name = "80th Percentile",
                    x = ~c(q80b, q80b), y = ~ c(0, max(fatalityb$y)),
                    mode = 'lines', type = 'scatter',
                    line = list(color = vir_col(3)[3]),
                    text = ~paste('Annual FWS estimate<br>survey + prior =',
                                  round(q80b, 2)),
                    hoverinfo = 'text'
                    )%>%
          add_trace(name = "Survey only",
                    x = ~fatality2b$x, y = ~fatality2b$y,
                    type = "scatter", 
                    mode = "lines",
                    fill = "tozeroy",
                    line = list(color = vir_col(3)[2]), 
                    text = ~paste("Probability of ",
                                  round(fatality2b$x, 2),
                                  " fatalities<br>using survey data = ",
                                  round(fatality2b$y, 2),
                                  sep = ""),
                    hoverinfo = "text")%>%
          add_trace(name = "80th Percentile",
                    x = ~c(q82b, q82b), y = ~ c(0, max(fatality2b$y)),
                    mode = 'lines', type = 'scatter',
                    line = list(color = vir_col(3)[2]),
                    text = ~paste('Annual FWS estimate<br>survey only =',
                                  round(q80b, 2)),
                    hoverinfo = 'text'                   
                    )%>%
          layout(##title = "Predicted Annual Eagle Fatalities",
                 xaxis = list(title = "Fatalities per Year"),
                 yaxis = list(title = "Probability Density"),
                 legend = list(x = 0.7,
                               y = 1))
    })
output$textb <-renderText({
    paste("For the proposed project observing ", input$min, "minutes of bald eagle flight time during ", input$hrs, " survey hours covering ", input$area, "(ha), the predicted take requiring mitigation is ", round(q80b, 1), " bald eagles using the update priors.")
    })
  #}
})
plotlyOutput('fatalityb')
```
